---
title: "data_cleaning"
author: "Aspen Coyle"
date: "2024-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Info

This script cleans and organizes police data on reports from Denny-Blaine and other beaches, with the goal of examining police response frequency over time

#### Load libraries, and install if necessary

```{r}
# Add all required libraries here
list.of.packages <- c("tidyverse", "lubridate")
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})

```


# Download data, remove unnecessary columns
```{r}
# Download data
db_call_dat <- read.csv("../data/db_call_data_2018-2023.csv")

# Examine some columns to see if we can drop

table(db_call_dat$Precinct)
# Precinct: all are east, safe to drop

table(db_call_dat$Sector)
# Sector: all are Charlie, safe to drop

table(db_call_dat$Beat)
# Beat: all are C3, safe to drop

# Can also drop Arrived Time (as non-arrived calls are blank, and we care about the date of the call over a long timeframe, not the response time)

db_call_dat <- db_call_dat %>%
  select(-c("Precinct", "Sector", "Beat", "Arrived.Time"))
```

# Explore and clean data

```{r}
# A small thing I noticed, there's a fair amount of untidy data, especially in the Final.Call.Type column.
# Using untidy data, SPD! For shame!!! 
# Let's clean that up a touch so R doesn't stumble
db_call_dat$Final.Call.Type <- str_remove(db_call_dat$Final.Call.Type, "--")


# Look at different reasons calls were made
table(db_call_dat$Initial.Call.Type)

# Alright, lots of these are non-applicable to any disorder or misconduct at Denny-Blaine.
# Let's remove all such non-applicable entries

# Remove rows with the following response types
db_call_dat <- db_call_dat %>%
  filter(Initial.Call.Type != "ALARM - COMM, SILENT/AUD BURG (INCL BANKS)") %>%         # burglary alarm
  filter(Initial.Call.Type != "ALARM - RESIDENTIAL - BURGLARY, SILENT/AUDIBLE") %>%     # burglary alarm
  filter(Initial.Call.Type != "ALARM - DURESS/PANIC,BUS/TAXI/CAR/PRSN - NOT DV") %>%    # car alarm
  filter(Initial.Call.Type != "ASLT - DV CRITICAL") %>%                                 # domestic violence
  filter(Initial.Call.Type != "ASSIST SPD") %>%                                         # assisting with other event
  filter(Initial.Call.Type != "AUTO - THEFT/LOSS OF PLATES") %>%                        # auto theft
  filter(Initial.Call.Type != "AUTO RECOVERY") %>%                                      # auto recovery
  filter(Initial.Call.Type != "AUTO THEFT - VEH THEFT OR THEFT AND RECOVERY") %>%       # auto theft/recovery
  filter(Initial.Call.Type != "BURG - OCCUPIED RESD") %>%                               # burglary
  filter(Initial.Call.Type != "BURG - RESD (INCL UNOCC STRUCTURES ON PROP)") %>%        # burglary
  filter(Initial.Call.Type != "CHILD - CRITICAL ABAND, ABUSE, MOLEST, NEGLECT") %>%     # child abuse
  filter(Initial.Call.Type != "DISTURBANCE - DV - NO ASLT") %>%                         # domestic violence
  filter(Initial.Call.Type != "DISTURBANCE - DV CRITICAL") %>%                          # domestic violence
  filter(Initial.Call.Type != "DOWN - CHECK FOR DOWN PERSON") %>%                       # wellness check
  filter(Initial.Call.Type != "DUI - DRIVING UNDER INFLUENCE") %>%                      # DUI
  filter(Initial.Call.Type != "FOLLOW UP") %>%                                          # follow up on incident
  filter(Initial.Call.Type != "FRAUD") %>%                                              # fraud
  filter(Initial.Call.Type != "HARBOR - WATER DEBRIS, NAVIGATIONAL HAZARDS") %>%        # water debris 
  filter(Initial.Call.Type != "HARBOR - WATER EMERGENCIES") %>%                         # water emergency
  filter(Initial.Call.Type != "MVC - HIT AND RUN") %>%                                  # hit and run
  filter(Initial.Call.Type != "OBS - MVC - NON INJURY, BLOCKING") %>%                   # motor vehicle collision
  filter(Initial.Call.Type != "OBS - MVC - UNK INJURIES") %>%                           # motor vehicle collision
  filter(Initial.Call.Type != "OBS - MVC - WITH INJURIES (INCLUDES HIT AND RUN)") %>%   # motor vehicle collision
  filter(Initial.Call.Type != "OBS - REPORT, NON INJ/NON BLKG OR AFTER FACT INJ") %>%   # one entry - checked, was MVC
  filter(Initial.Call.Type != "ORDER  - SERVICE OF DV COURT ORDER") %>%                 # serving court order
  filter(Initial.Call.Type != "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY") %>%       # premise check
  filter(Initial.Call.Type != "PROPERTY - DAMAGE (GUN RELATED)") %>%                    # gun related property damage
  filter(Initial.Call.Type != "SERVICE - WELFARE CHECK") %>%                            # welfare check
  filter(Initial.Call.Type != "SFD - ASSIST ON CRITICAL FIRE OR MEDIC RESPONSE") %>%    # fire or medic response
  filter(Initial.Call.Type != "SHOTS HEARD - DELAY") %>%                                # gun violence
  filter(Initial.Call.Type != "SHOTS HEARD - IN PROGRESS") %>%                          # gun violence
  filter(Initial.Call.Type != "SUICIDE - CRITICAL") %>%                                 # suicide
  filter(Initial.Call.Type != "SUICIDE - SUICIDAL PERSON AND ATTEMPTS") %>%             # suicide
  filter(Initial.Call.Type != "TRAFFIC - MOVING VIOLATION") %>%                         # traffic stop
  filter(Initial.Call.Type != "TRAFFIC - ROAD RAGE") %>%                                # road rage
  filter(Initial.Call.Type != "TRAFFIC STOP - OFFICER INITIATED ONVIEW") %>%            # traffic stop
  filter(Initial.Call.Type != "TREES DOWN, OBSTRUCTING PUBLIC PROP (NO HAZ)") %>%       # fallen trees
  filter(Initial.Call.Type != "TRU - FOLLOW UP") %>%                                    # follow up
  filter(Initial.Call.Type != "TRU - FRAUD") %>%                                        # fraud
  filter(Initial.Call.Type != "TRU - MVC - HIT AND RUN") %>%                            # motor vehicle collision
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - LANDLINE (INCLUDES OPEN LINE)") %>%  # one entry, residential alarm
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - WRLS PHNS (INCL OPEN LINE)") %>%     # unknown complaint
  filter(Initial.Call.Type != "UNKNOWN - COMPLAINT OF UNKNOWN NATURE")                  # two, both were DV

# Alright, we've got 91 reports that could potentially have something vaguely to do with DB, at least according to the initial report! Let's look at the final call reports to see if we can rule anything out

table(db_call_dat$Final.Call.Type)

# Remove ones that can't be DB
db_call_dat <- db_call_dat %>%
  filter(Final.Call.Type != "BURGLARY - RESIDENTIAL OCCUPIED") %>%              # burglary, was called for disturbance
  filter(Final.Call.Type != "BURGLARY - UNOCC STRUC ON RESN PROP") %>%          # burglary, was called for disturbance
  filter(Final.Call.Type != "CASUALTY,NON-TRAF,NON-CRIM - NON-DRUG RELATED")    # dead body, non-drug-related

# Let's also do a quick check of event resolution

table(db_call_dat$Event.Clearance.Description)

# Remove false report
db_call_dat <- db_call_dat %>%
  filter(Final.Call.Type != "FALSE COMPLAINT/UNFOUNDED")
```

Cool, we've got a list of incidents that could have potentially been at DB

Let's sort these into a few neater categories - property damage/theft, lewd conduct, disturbance, drugs, traffic/parking, suspicious people

# Organizing into broad categories

```{r}
# Add new column, fill with NAs
db_call_dat$broad_category <- NA

# We'll do this in base R - sorry, I just prefer this way!

# do based on initial call type, since those have more precision
table(db_call_dat$Initial.Call.Type)

db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ALARM - DURESS/PANIC,TAXI/CAR/PRSN - NOT DV"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "BURN - RECKLESS BURNING"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "DISTURBANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "HARAS - NO BIAS OR THREATS"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "LEWD -  (EXPOSING, FLASHING)"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "NUISANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - DOA - CASUALTY, DEAD BODY"] <- "drugs"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - FIGHT - VERBAL/ORAL (NO WEAPONS)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - PROWLER"] <- "susp.pers"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OVERDOSE"] <- "drugs"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PARKING VIOLATION (EXCEPT ABANDONED CAR)"] <- "traffic"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PROPERTY - DAMAGE"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SEX IN PUBLIC PLACE (INCL MASTURBATION)"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SUSPICIOUS PERSON, VEHICLE, OR INCIDENT"] <- "susp.pers"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SUSPICIOUS STOP - OFFICER INITIATED ONVIEW"] <- "susp.pers"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "THEFT (DOES NOT INCLUDE SHOPLIFT OR SVCS)"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "THREATS (INCLS IN-PERSON/BY PHONE/IN WRITING)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRAFFIC - BLOCKING ROADWAY"] <- "traffic"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRESPASS"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - DISTURBANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - PROPERTY DESTRUCTION/DAMAGE"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - SEX IN PUBLIC"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - SUSPICIOUS CIRCUMSTANCES"] <- "susp.pers"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - THEFT"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "URINATING, DEFECATING IN PUBLIC"] <- "lewd"
```

Nice, now we've got a slick clean (and TIDY, please take notes SPD!) dataframe. Now let's pull out the year from the date/time info!

```{r}
# Let's split the time/date string and just select the date
db_call_dat$Original.Time.Queued <- str_split_i(db_call_dat$Original.Time.Queued, " ", 1)

# Rename column
db_call_dat <- db_call_dat %>%
  rename(Incident.Date = Original.Time.Queued)

# Check column type
str(db_call_dat$Incident.Date)
# Yep it's a character string

# Transform into Date object type
db_call_dat$Incident.Date <- mdy(db_call_dat$Incident.Date)
# Check it worked
str(db_call_dat$Incident.Date)

# Write everything out
write.csv(db_call_dat, file = "../output/potential_db_calls_2018-2023.csv", row.names = FALSE)
```
Alright, next step: creating visualizations looking at trends by year!



