---
title: "data_cleaning"
author: "Aspen Coyle"
date: "2024-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Info

This script cleans and organizes police data on reports from Denny-Blaine and other beaches, with the goal of examining police response frequency over time

#### Load libraries, and install if necessary

```{r}
# Add all required libraries here
list.of.packages <- c("tidyverse", "lubridate")
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})

```


# Download data, remove unnecessary columns

Data was downloaded from the SPD public call data page, available at https://data.seattle.gov/Public-Safety/Call-Data/33kz-ixgy

Downloaded all available data from East precinct, Charlie sector on 2024-05-18




```{r}
# Read in data
db_call_dat <- read.csv("../data/db_call_data_2010-2023.csv")

# Woof, 237,000 rows! 

# Let's narrow it down some, and only select those calls with a blurred lat/long within a quarter mile of Denny Blaine
# North boundary: 47.623689
# South boundary: 47.616472
# East boundary: -122.274502
# West boundary: -122.285688

# Filter by location
db_call_dat <- db_call_dat %>%
  dplyr::filter(Blurred_Latitude < 47.623689 & Blurred_Latitude > 47.616472) %>%    # north/south
  dplyr::filter(Blurred_Longitude < -122.274502 & Blurred_Longitude > -122.285688)  # east/west 


# Dropping unnecessary columns
# Precinct: all are East, safe to drop
# Sector: all are Charlie, safe to drop
# Beat: let's check
table(db_call_dat$Beat)
# All are C3. Regardless, safe to drop

# Can also drop Arrived Time (as non-arrived calls are blank, and we care about the date of the call over a long timeframe, not the response time)

db_call_dat <- db_call_dat %>%
  select(-c("Precinct", "Sector", "Beat", "Arrived.Time"))
```

# Explore and clean data

```{r}
# A small thing I noticed, there's a fair amount of untidy data, especially in the Final.Call.Type column.
# Using untidy data, SPD! For shame!!! 
# Let's clean that up a touch so R doesn't stumble
db_call_dat$Final.Call.Type <- str_remove(db_call_dat$Final.Call.Type, "--")


# Look at different reasons calls were made
table(db_call_dat$Initial.Call.Type)

# Alright, lots of these are non-applicable to any disorder or misconduct at Denny-Blaine.
# Let's remove all such non-applicable entries

# Remove rows with the following response types
db_call_dat <- db_call_dat %>%
  filter(Initial.Call.Type != "-ASSIGNED DUTY - FOOT BEAT (FROM ASSIGNED CAR)") %>%    # patrol
  filter(Initial.Call.Type != "ALARM - COMM, HOLD-UP/PANIC (EXCEPT BANKS)") %>%        # burglary
  filter(Initial.Call.Type != "ALARM - COMM, SILENT/AUD BURG (INCL BANKS)") %>%        # burglary
  filter(Initial.Call.Type != "ALARM - DURESS/PANIC,BUS/TAXI/CAR/PRSN - NOT DV") %>%   # burglary (one instance, checked)
  filter(Initial.Call.Type != "ALARM - RESIDENTIAL - BURGLARY, SILENT/AUDIBLE") %>%    # burglary
  filter(Initial.Call.Type != "ALARM - RESIDENTIAL - SILENT/AUD PANIC/DURESS") %>%     # burglary
  filter(Initial.Call.Type != "ASLT - DV CRITICAL") %>%                                # domestic violence
  filter(Initial.Call.Type != "ASSIST OTHER AGENCY - ROUTINE SERVICE") %>%             # assist other agency
  filter(Initial.Call.Type != "ASSIST PUBLIC - NO WELFARE CHK OR DV ORDER SERVICE") %>%# assist public
  filter(Initial.Call.Type != "ASSIST SPD") %>%                                        # assist SPD
  filter(Initial.Call.Type != "AUTO RECOVERY") %>%                                     # auto recovery
  filter(Initial.Call.Type != "BURG - COMM BURGLARY") %>%                              # burglary
  filter(Initial.Call.Type != "BURG - OCCUPIED RESD") %>%                              # burglary
  filter(Initial.Call.Type != "BURG - RESD (INCL UNOCC STRUCTURES ON PROP)") %>%       # burglary
  filter(Initial.Call.Type != "CHILD - ABAND, ABUSE, MOLEST, NEGLECT") %>%             # child abuse
  filter(Initial.Call.Type != "CHILD - CRITICAL ABAND, ABUSE, MOLEST, NEGLECT") %>%    # child abuse
  filter(Initial.Call.Type != "DISTURBANCE - DV - NO ASLT") %>%                        # domestic violence
  filter(Initial.Call.Type != "DISTURBANCE - DV CRITICAL") %>%                         # domestic violence
  filter(Initial.Call.Type != "DOWN - CHECK FOR DOWN PERSON") %>%                      # welfare check
  filter(Initial.Call.Type != "DUI - DRIVING UNDER INFLUENCE") %>%                     # DUI
  filter(Initial.Call.Type != "FOLLOW UP") %>%                                         # follow up
  filter(Initial.Call.Type != "FOUND PERSON") %>%                                      # found person
  filter(Initial.Call.Type != "FRAUD") %>%                                             # fraud
  filter(Initial.Call.Type != "HARBOR - WATER DEBRIS, NAVIGATIONAL HAZARDS") %>%       # water hazard
  filter(Initial.Call.Type != "HARBOR - WATER EMERGENCIES") %>%                        # water emergencies
  filter(Initial.Call.Type != "HARBOR - WATER NON-EMERGENCIES") %>%                    # water non-emergencies
  filter(Initial.Call.Type != "INFORMATIONAL BROADCASTS") %>%                          # informational broadcasts
  filter(Initial.Call.Type != "JUVENILE - RUNAWAY") %>%                                # juvenile runaway
  filter(Initial.Call.Type != "JUVENILE - RUNAWAY PICKUP") %>%                         # runaway found
  filter(Initial.Call.Type != "MISSING - ADULT") %>%                                   # adult missing
  filter(Initial.Call.Type != "MISSING - CHILD") %>%                                   # child missing
  filter(Initial.Call.Type != "MVC - HIT AND RUN") %>%                                 # motor vehicle collision
  filter(Initial.Call.Type != "OBS - ACC - UNK INJURIES") %>%                          # checked, all traffic accidents
  filter(Initial.Call.Type != "OBS - ACC - WITH INJURIES (INCLUDES HIT AND RUN)") %>%  # motor vehicle collision
  filter(Initial.Call.Type != "OBS - BURG - IP/JO - COMM BURG (INCLUDES SCHOOLS)") %>% # burglary
  filter(Initial.Call.Type != "OBS - DOWN - CHECK FOR PERSON DOWN") %>%                # welfare check
  filter(Initial.Call.Type != "OBS - HIT AND RUN (NO INJURIES). INCLUDES IP/JO") %>%   # motor vehicle collision
  filter(Initial.Call.Type != "OBS - INJURED - IP/JO - PERSON/INDUSTRIAL ACCIDENT") %>%# accident
  filter(Initial.Call.Type != "OBS - MVC - NON INJURY, BLOCKING") %>%                  # motor vehicle collision
  filter(Initial.Call.Type != "OBS - MVC - UNK INJURIES") %>%                          # motor vehicle collision
  filter(Initial.Call.Type != "OBS - MVC - WITH INJURIES (INCLUDES HIT AND RUN)") %>%  # motor vehicle collision
  filter(Initial.Call.Type != "OBS - NON INJURY, BLOCKING") %>%                       # checked, all MVCs
  filter(Initial.Call.Type != "OBS - REPORT, NON INJ/NON BLKG OR AFTER FACT INJ") %>%  # motor vehicle collision
  filter(Initial.Call.Type != "OBS - SECONDARY - FORGERY/BUNCO/SCAMS/ID THEFT") %>%    # fraud
  filter(Initial.Call.Type != "OBS - TRU - ACC - HIT AND RUN") %>%                     # motor vehicle collision
  filter(Initial.Call.Type != "OPEN - BUILDING, DOOR, ETC.") %>%                       # open building
  filter(Initial.Call.Type != "ORDER - CRITICAL VIOLATION OF DV COURT ORDER") %>%      # court order violation
  filter(Initial.Call.Type != "ORDER - VIOLATING DV COURT ORDER") %>%                  # court order violation
  filter(Initial.Call.Type != "ORDER - VIOLATION OF COURT ORDER (NON DV)") %>%         # court order violation
  filter(Initial.Call.Type != "ORDER  - SERVICE OF DV COURT ORDER") %>%                # service of court order
  filter(Initial.Call.Type != "PEACE-STANDBY TO ASSURE (NO COURT ORDR SVC)") %>%       # checked, public assistance and DV
  filter(Initial.Call.Type != "POWER OUT - POLES AND TRANSFORMERS") %>%                # power outage
  filter(Initial.Call.Type != "PREDICTIVE POLICING") %>%                               # predictive policing
  filter(Initial.Call.Type != "PROPERTY - FOUND") %>%                                  # found property
  filter(Initial.Call.Type != "PROPERTY - FOUND GUN, SHELLCASINGS") %>%                # found property
  filter(Initial.Call.Type != "SERVICE - WELFARE CHECK") %>%                           # welfare check
  filter(Initial.Call.Type != "SFD - ASSIST ON CRITICAL FIRE OR MEDIC RESPONSE") %>%   # assisting fire or medics
  filter(Initial.Call.Type != "SUICIDE - CRITICAL") %>%                                # suicide
  filter(Initial.Call.Type != "SUICIDE - SUICIDAL PERSON AND ATTEMPTS") %>%            # suicide
  filter(Initial.Call.Type != "THEFT OF SERVICES") %>%                                 # theft of services
  filter(Initial.Call.Type != "TRAF - ELUDING POLICE") %>%                             # eluding police
  filter(Initial.Call.Type != "TRAFFIC - ASSIST MOTORIST") %>%                         # assisting motorist
  filter(Initial.Call.Type != "TRAFFIC - BO SIGNALS AND DOWN SIGNS") %>%               # traffic signals down
  filter(Initial.Call.Type != "TRAFFIC - MOVING VIOLATION") %>%                        # moving violation
  filter(Initial.Call.Type != "TRAFFIC - ROAD RAGE") %>%                               # road rage
  filter(Initial.Call.Type != "TRAFFIC STOP - OFFICER INITIATED ONVIEW") %>%           # traffic stop
  filter(Initial.Call.Type != "TREES DOWN, OBSTRUCTING PUBLIC PROP (NO HAZ)") %>%      # downed trees
  filter(Initial.Call.Type != "TRU - FOLLOW UP") %>%                                   # follow up
  filter(Initial.Call.Type != "TRU - FRAUD") %>%                                       # fraud
  filter(Initial.Call.Type != "TRU - MVC - HIT AND RUN") %>%                           # motor vehicle collision
  filter(Initial.Call.Type != "WATER FLOODS (BROKEN MAINS/HYDRANTS, NO HAZ)")          # water floods



# Alright, we've got some reports that could potentially have something vaguely to do with DB, at least according to the initial report! Let's look at the final call reports to see if we can rule anything out

table(db_call_dat$Final.Call.Type)

# Remove ones that can't be DB
db_call_dat <- db_call_dat %>%
  filter(Final.Call.Type != "ALARM-COMM (INC BANK, ATM, SCHOOLS, BSN)") %>%              # burglary
  filter(Final.Call.Type != "ALARM-RESIDENTIAL BURG") %>%                                # burglary
  filter(Final.Call.Type != "ALARM-RESIDENTIAL PANIC OR DURESS") %>%                     # residential assistance
  filter(Final.Call.Type != "AUTOMOBILES - RECOVERY (THEFT)") %>%                        # recovered car
  filter(Final.Call.Type != "BURGLARY - RESIDENTIAL OCCUPIED") %>%                       # burglary
  filter(Final.Call.Type != "BURGLARY - RESIDENTIAL, UNOCCUPIED") %>%                    # burglary
  filter(Final.Call.Type != "BURGLARY - UNOCC STRUC ON RESN PROP") %>%                   # burglary
  filter(Final.Call.Type != "CASUALTY,NON-TRAF,NON-CRIM - NON-DRUG RELATED") %>%         # non-drug-related casualty
  filter(Final.Call.Type != "DIRECTED PATROL ACTIVITY") %>%                              # patrol
  filter(Final.Call.Type != "DISTURBANCE - NOISE RESIDENTIAL") %>%                       # residential noise
  filter(Final.Call.Type != "DV - ARGUMENTS, DISTURBANCE (NO ARREST)") %>%               # domestic violence
  filter(Final.Call.Type != "DV - DOMESTIC THREATS BY PHONE OR WRITING") %>%             # domestic violence
  filter(Final.Call.Type != "DV - DOMESTIC VIOL/ASLT (ARREST MANDATORY)") %>%            # domestic violence
  filter(Final.Call.Type != "FRAUD - FRAUD, INCLUDING BUNCO") %>%                        # fraud
  filter(Final.Call.Type != "HARBOR - WATER EMERGENCIES") %>%                            # water emergency
  filter(Final.Call.Type != "HAZARDS - HAZARDS") %>%                                     # general hazard
  filter(Final.Call.Type != "HELP THE OFFICER -ASSIST THE OFFICER (NON EMERG)") %>%      # assisting officer
  filter(Final.Call.Type != "OBS - DOA - CASUALTY, DEAD BODY") %>%                       # dead body
  filter(Final.Call.Type != "PERSON - RUNAWAY") %>%                                      # runaway
  filter(Final.Call.Type != "PREDICTIVE POLICING PATROL") %>%                            # predictive policing
  filter(Final.Call.Type != "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY") %>%          # premise check
  filter(Final.Call.Type != "PREMISE CHECKS - BUSINESS CHECK") %>%                       # premise check
  filter(Final.Call.Type != "PREMISE CHECKS - HOMELESS ENCAMPMENT") %>%                  # premise check
  filter(Final.Call.Type != "PREMISE CHECKS - REQUEST TO WATCH") %>%                     # premise check
  filter(Final.Call.Type != "SUSPICIOUS CIRCUM. - BUILDING (OPEN DOOR, ETC)") %>%        # building 
  filter(Final.Call.Type != "SUSPICIOUS CIRCUM. - SUSPICIOUS VEHICLE") %>%               # suspicious vehicle
  filter(Final.Call.Type != "THEFT - SHOPLIFT") %>%                                      # shoplifting
  filter(Final.Call.Type != "TRAFFIC - D.U.I") %>%                                       # DUI
  filter(Final.Call.Type != "TRAFFIC - MOVING VIOLATION") %>%                            # moving violation
  filter(Final.Call.Type != "TRAFFIC - MV COLLISION INVESTIGATION") %>%                  # motor vehicle collision
  filter(Final.Call.Type != "WARRANT SERVICES - MISDEMEANOR")                            # warrant services

# Let's also do a quick check of event resolution

table(db_call_dat$Event.Clearance.Description)

# Remove false report
db_call_dat <- db_call_dat %>%
  filter(Final.Call.Type != "FALSE COMPLAINT/UNFOUNDED")

# Alright, 1300-odd rows. Let's go back through, this time removing combinations that don't work
db_call_dat <- db_call_dat %>%
  filter(Initial.Call.Type != "DIRECTED PATROL ACTIVITY" | Final.Call.Type != "PREMISE CHECKS - CRIME PREVENTION") %>%                 # checking premises
  filter(Initial.Call.Type != "DISTURBANCE" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)") %>%                          # public assistance
  filter(Initial.Call.Type != "HAZ - POTENTIAL THRT TO PHYS SAFETY" | Final.Call.Type != "ASSIST OTHER AGENCY - CITY AGENCY") %>%      # assisting agency
  filter(Initial.Call.Type != "HAZ - POTENTIAL THRT TO PHYS SAFETY" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)") %>%  # public assistance
  filter(Initial.Call.Type != "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY" | Final.Call.Type != "ASSIST OTHER AGENCY - CITY AGENCY") %>%     # assisting agency
  filter(Initial.Call.Type != "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)") %>% # assisting agency
  filter(Initial.Call.Type != "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY" | Final.Call.Type != "PREMISE CHECKS - CRIME PREVENTION") %>%     # premise checks
  filter(Initial.Call.Type != "REQUEST TO WATCH" | Final.Call.Type != "PREMISE CHECKS - CRIME PREVENTION") %>%                                 # premise checks
  filter(Initial.Call.Type != "SERVICE - WELFARE CHECK" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)") %>%                      # welfare check
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - LANDLINE (INCLUDES OPEN LINE)" | Final.Call.Type != "ASSIST OTHER AGENCY - CITY AGENCY") %>%# assisting agency
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - LANDLINE (INCLUDES OPEN LINE)" | Final.Call.Type != "ASSIST PUBLIC - 911 HANG UP, OPEN LINE") %>%  # 911 hangup
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - LANDLINE (INCLUDES OPEN LINE)" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)") %>% # assisting public
  filter(Initial.Call.Type != "UNKNOWN - ANI/ALI - WRLS PHNS (INCL OPEN LINE)" | Final.Call.Type != "ASSIST OTHER AGENCY - CITY AGENCY") %>%   # assisting agency
  filter(Initial.Call.Type != "UNKNOWN - COMPLAINT OF UNKNOWN NATURE" | Final.Call.Type != "ASSIST OTHER AGENCY - CITY AGENCY") %>%            # assisting agency
  filter(Initial.Call.Type != "UNKNOWN - COMPLAINT OF UNKNOWN NATURE" | Final.Call.Type != "ASSIST PUBLIC - OTHER (NON-SPECIFIED)")            # assisting public
```

Cool, we've got a list of incidents that could have potentially been at DB

Let's sort these into a few neater categories (abbreviations in parentheses):
- Disturbance
- Drugs
- Lewd conduct (lewd)
- Property damage/theft (prop.dmg)
- Suspicious people (suspicious)
- Violence
# Organizing into broad categories

```{r}
# Add new column, fill with NAs
db_call_dat$broad_category <- NA

# We'll do this in base R - sorry, I just prefer this way!

# do based on initial call type. This is potentially flawed, as assumes initial call was accurate representation of situation, which is often VERY not true. If anything, this dramatically errs on the high side
table(db_call_dat$Initial.Call.Type)

# Start assigning
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ALARM - DURESS/PANIC,TAXI/CAR/PRSN - NOT DV"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ANIMAL - INJURED, DEAD HAZARD, OTHER"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ASLT - CRITICAL (NO SHOOTINGS)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ASLT - REPORT FOR PERSON SHOT OR SHOT AT"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ASLT - WITH OR W/O WPNS (NO SHOOTINGS)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "AUTO - THEFT/LOSS OF PLATES"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "AUTO THEFT - VEH THEFT OR THEFT AND RECOVERY"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "BIAS"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "BURN - RECKLESS BURNING"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "DIRECTED PATROL ACTIVITY"] <- "disturbance" # one entry, was liquor viol
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "DISTURBANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "EXPLOSION"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "FIGHT - CRITICAL WITH WEAPONS"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "FIGHT - PHYSICAL (NO WEAPONS)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "FIREWORKS (NO HAZARD)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "HARAS - NO BIAS OR THREATS"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "HAZ - POTENTIAL THRT TO PHYS SAFETY"] <- "suspicious"   # checked this
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ILLEGAL DUMPING"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "LEWD -  (EXPOSING, FLASHING)"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "LIQUOR VIOLATIONS - ADULT"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "LIQUOR VIOLATIONS - MINOR"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "NARCOTICS - VIOLATIONS (LOITER, USE, SELL, NARS)"] <- "drug"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "NOISE - ANIMAL (INCLUDES BARKING DOGS)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "NOISE - GENERAL"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "NUISANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - ANIMAL - IP/JO - BITE"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - ARSON - IP/JO"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - DOA - CASUALTY, DEAD BODY"] <- "drug"      # was OD
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - EXPLOSION - IP/JO"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - FIGHT - IP - PHYSICAL (NO WEAPONS)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - FIGHT - VERBAL/ORAL (NO WEAPONS)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - MENTAL - PERSON OR PICKUP/TRANSPORT"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - NOISE - DISTURBANCE (PARTY, ETC)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - PROWLER"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - PROWLER - IP/JO"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - SECONDARY - PROPERTY DAMAGE/DESTRUCTION"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS - SECONDARY - THEFT"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OBS WEAPN-IP/JO-GUN,DEADLY WPN (NO THRT/ASLT/DIST)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "OVERDOSE"] <- "drug"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PARKING VIOLATION (EXCEPT ABANDONED CAR)"] <- "traffic"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PARKS VIOLATIONS/CITES (INCLUDES EXCLUSIONS)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PERSON IN BEHAVIORAL/EMOTIONAL CRISIS"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PREMISE CHECK, OFFICER INITIATED ONVIEW ONLY"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PROPERTY - DAMAGE"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "PROPERTY - DAMAGE (GUN RELATED)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "REQUEST TO WATCH"] <- "suspicious"  # manually checked
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ROBBERY - CRITICAL (INCLUDES STRONG ARM)"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "ROBBERY (INCL STRONG ARM)"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SEX IN PUBLIC PLACE (INCL MASTURBATION)"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SHOTS HEARD - DELAY"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SHOTS HEARD - IN PROGRESS"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SUSPICIOUS PERSON, VEHICLE, OR INCIDENT"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "SUSPICIOUS STOP - OFFICER INITIATED ONVIEW"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "THEFT (DOES NOT INCLUDE SHOPLIFT OR SVCS)"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "THREATS (INCLS IN-PERSON/BY PHONE/IN WRITING)"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRAFFIC - BLOCKING ROADWAY"] <- "traffic"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRESPASS"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - ASLT - MOLESTED ADULT (GROPED, FONDLED, ETC)"] <- "violence"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - DISTURBANCE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - HARASSMENT"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - PROPERTY DESTRUCTION/DAMAGE"] <- "prop.dmg"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - SEX IN PUBLIC"] <- "lewd"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - SUSPICIOUS CIRCUMSTANCES"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "TRU - THEFT"] <- "theft"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "UNKNOWN - ANI/ALI - LANDLINE (INCLUDES OPEN LINE)"] <- "suspicious"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "UNKNOWN - COMPLAINT OF UNKNOWN NATURE"] <- "disturbance"
db_call_dat$broad_category[db_call_dat$Initial.Call.Type == "WEAPN - GUN,DEADLY WPN (NO THRTS/ASLT/DIST)"] <- "violence"

# Make sure we had minimal typos
table(db_call_dat$broad_category)
```

Nice, now we've got a slick clean (and TIDY, please take notes SPD!) dataframe. Now let's pull out the year from the date/time info!

```{r}
# Let's split the time/date string and just select the date
db_call_dat$Original.Time.Queued <- str_split_i(db_call_dat$Original.Time.Queued, " ", 1)

# Rename column
db_call_dat <- db_call_dat %>%
  rename(Incident.Date = Original.Time.Queued)

# Check column type
str(db_call_dat$Incident.Date)
# Yep it's a character string

# Transform into Date object type
db_call_dat$Incident.Date <- mdy(db_call_dat$Incident.Date)
# Check it worked
str(db_call_dat$Incident.Date)

# Write everything out
write.csv(db_call_dat, file = "../output/potential_db_calls_2010-2023.csv", row.names = FALSE)
```
Alright, next step: creating visualizations looking at trends by year!



