---
title: "Data Visualization"
author: "Aspen Coyle"
date: "2024-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script visualizes the data cleaned in the previous script

#### Load libraries, and install if necessary

```{r}
# Add all required libraries here
list.of.packages <- c("tidyverse", "lubridate", "viridisLite", "plotly")
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})
```


# Visualizations

```{r}
# Read in data
db_call_dat <- read.csv(file = "../output/potential_db_calls_2010-2023.csv")
seattle_pop <- read.csv(file = "../data/seattle_pop_data_2010-2023.csv")

# Create new Year column
db_call_dat$Year <- isoyear(db_call_dat$Incident.Date)

# Remove all data from the year 2024, as it's only just beginning :)
db_call_dat <- db_call_dat %>%
  filter(Year != 2024)

# Create line plot of broad call type vs. year

# Mutate data
graph_dat <- db_call_dat %>%
  group_by(broad_category, Year, .drop = FALSE) %>%
  summarise(n = length(broad_category))

# Convert to tibble
graph_dat <- as_tibble(graph_dat)

# Add in zeros
graph_dat <- graph_dat %>% 
  complete(broad_category, Year, fill = list(n = 0))

# Join population data
graph_dat <- left_join(x = graph_dat, y = seattle_pop, by = "Year")

# Change population to int
graph_dat$Population <- as.integer(graph_dat$Population)

# Calculate rate per 100,000 people
graph_dat$n_rate <- (graph_dat$n/graph_dat$Population)*100000
```


# Create graphs

```{r}
# Create graph
graph <- graph_dat %>%
  ggplot(aes(x = Year, y = n, group = broad_category, color = broad_category)) +
  geom_line(size = 1) +
  ylim(0, 40) +
  labs(y = "Number of Calls", col = "Call Type") +
  ggtitle("Calls to SPD Potentially Related to Denny-Blaine")
  
# Save graph
ggsave(filename = "../graphs/DennyBlaine/potential_DB_calls_2010-2023_linegraph.png", plot = graph,
       width = 10, height = 7, units = "in")

# Create string of labels
labels <- c(
  disturbance = "Disturbance", 
  drug = "Drug-Related", 
  lewd = "Lewd Conduct", 
  prop.dmg = "Property Damage", 
  suspicious = "Suspicious Behavior", 
  theft = "Theft", 
  traffic = "Parking and Traffic", 
  violence = "Violence")

# Raw Numbers, Un-Normalized for Population
graph <- graph_dat %>%
  ggplot(aes(x = Year, y = n, fill = broad_category)) +
  geom_area() +
  scale_fill_viridis_d() +
  theme(legend.position = "none", 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 8)) +
  facet_wrap(~broad_category, scales = "fixed", labeller = labeller(broad_category = labels)) +
  ggtitle("Calls to SPD Potentially Related to Denny-Blaine (2010-2023)") +
  ylab("Number of Calls")

ggsave(filename = "../graphs/DennyBlaine/potential_db_calls_2010-2023_facetwrap.png", plot = graph,
       width = 10, height = 7, units = "in")

# Population-Normalized Calls
graph <- graph_dat %>%
  ggplot(aes(x = Year, y = n_rate, fill = broad_category)) +
  geom_area() +
  scale_fill_viridis_d() +
  theme(legend.position = "none", 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 8)) +
  facet_wrap(~broad_category, scales = "fixed", labeller = labeller(broad_category = labels)) +
  ggtitle("Calls to SPD Potentially Related to Denny-Blaine (2010-2023) Per 100,000 Seattle Residents") +
  ylab("Number of Calls Per 100,000 Seattleites")

ggsave(filename = "../graphs/DennyBlaine/pop_normalized_potential_db_calls_2010-2023_facetwrap.png", plot = graph,
       width = 10, height = 7, units = "in")
```

# Visualizing More Niche Calls

Alright, now it's time to look into specific calls relating to violations of the following municipal codes.
These are all named on the Denny Blaine for All website:

- 18.12.080: Animals running at large prohibited
- 18.12.145: Urinating or defecating prohibited except in restrooms
- 18.12.255: Liquor offenses
- 18.12.257: Unlawful possession of liquor in a public park
- 18.12.260: Littering

The other violations mentioned in the text on their website are either already captured (use of driveways and boulevards = traffic and parking), or can't really be examined with these data (general park operating hours, trespass in park)


```{r}
# ANIMALS

# Search df for rows containing the string "animal", create new column with a 1 if it has the string, a 0 if not
db_call_dat$Animals <- apply(db_call_dat, 1, function(x)as.integer(any(grep("animal", x, ignore.case = TRUE))))

# Number of animal-related incidents for all years
sum(db_call_dat$Animals)

# There's 5 the whole time! That's not worth graphing

# URINATING AND DEFECATING

# Search DF for rows containing the string "urin"
db_call_dat$Urine <- apply(db_call_dat, 1, function(x)as.integer(any(grep("urin", x, ignore.case = TRUE))))

# Number of urination and defecation incidents for all years
sum(db_call_dat$Urine)

# There's none!

# LIQUOR-RELATED OFFENSES

# Search DF for rows containing the string "liquor"
db_call_dat$Liquor <- apply(db_call_dat, 1, function(x)as.integer(any(grep("liquor", x, ignore.case = TRUE))))

# Number of offenses
sum(db_call_dat$Liquor)

# 11 total

# LITTER-RELATED OFFENSES

# Search DF for rows containing the string "litter"
db_call_dat$litter <- apply(db_call_dat, 1, function(x)as.integer(any(grep("litter", x, ignore.case = TRUE))))

# Number of offenses
sum(db_call_dat$litter)

# 0 total

```
Summary of calls made from 2010-2023: 
- Animal-related incidents: 5
- Urination and defecation: 0
- Liquor use and intoxication: 11
- Littering: 0

